---
title: "Data Visualisation Project"
author: "Prasad Kusumanti"
project: "olympics athlete" 
editor: visual
---
```{r, message=FALSE, warning=FALSE}
library(dplyr) 
```

```{r}
#| echo: false
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
olympics <- read_csv("olympics.csv", show_col_types = FALSE)
```

# Dataset

| Variable | Class | Description |
|----------|-------|-------------|
| `id` | double | Athlete ID |
| `name` | character | Athlete Name |
| `sex` | character | Athlete Sex |
| `age` | double | Athlete Age |
| `height` | double | Athlete Height in cm |
| `weight` | double | Athlete weight in kg |
| `team` | character | Country/Team competing for |
| `noc` | character | NOC region |
| `games` | character | Olympic games name |
| `year` | double | Year of Olympics |
| `season` | character | Season (either winter or summer) |
| `city` | character | City of Olympic host |
| `sport` | character | Sport |
| `event` | character | Specific event |
| `medal` | character | Medal (Gold, Silver, Bronze, or NA) |

## Athlete Participation in Selected Sports
```{r}
selected_sports <- c("Swimming", "Gymnastics", "Wrestling", "Football", "Hockey", "Boxing", "Badminton", "Archery")
filtered_data <- olympics %>% filter(sport %in% selected_sports)
ggplot(filtered_data, aes(y = reorder(sport, -table(sport)[sport]))) +
  geom_bar(fill = "blue") +
  labs(title = "Athlete Participation in Selected Sports", 
       x = "Number of Athletes", y = "Sport") +
  theme_minimal()
```
#Height vs Weight of Athletes by Sport
```{r}
library(dplyr)
library(ggplot2)
library(readr)
selected_sports <- c("Swimming", "Gymnastics", "Wrestling", "Football", "Hockey", "Boxing", "Badminton", "Archery")
filtered_data <- olympics %>%
  filter(sport %in% selected_sports) %>%
  filter(!is.na(height), !is.na(weight), height > 0, weight > 0)
ggplot(filtered_data, aes(x = height, y = weight, color = sport)) +
  geom_point(alpha = 0.6) +
  labs(
    title = "Height vs Weight of Athletes by Sport",
    x = "Height (cm)",
    y = "Weight (kg)",
    color = "Sport"
  ) +
  theme_minimal() +
  facet_wrap(~ sport, scales = "free")  
```
#Gender Representation in Selected Sports
```{r}
ggplot(filtered_data, aes(x = sport, fill = sex)) +
  geom_bar(position = "dodge") +
  labs(title = "Gender Representation in Selected Sports", 
       x = "Sport", y = "Number of Athletes", fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#Top Sports Played by Most Females
```{r}
female_sports <- olympics %>%
  filter(sex == "F") %>%
  count(sport, name = "female_count") %>%
  arrange(desc(female_count)) %>%
  head(15) %>%
  filter(sport != first(sport)) 

ggplot(female_sports, aes(x = reorder(sport, -female_count), y = female_count)) +
  geom_bar(stat = "identity", fill = "orange") + 
  labs(title = "Top Sports Played by Most Females", 
       x = "Sport", y = "Number of Female Athletes") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```
#Top Sports with Highest Medal-Winning Countries
```{r}
team_medals <- olympics %>%
  filter(!is.na(medal)) %>%
  group_by(noc, sport) %>%
  summarise(medal_count = n(), .groups = "drop") %>%
  arrange(sport, desc(medal_count))

top_teams_per_sport <- team_medals %>%
  group_by(sport) %>%
  slice_max(medal_count, n = 1) %>%
  ungroup()

top_teams_per_sport <- top_teams_per_sport %>%
  filter(sport != "Athletics") %>%
  arrange(desc(medal_count)) %>%
  head(10)

ggplot(top_teams_per_sport, aes(x = reorder(sport, -medal_count), y = medal_count, fill = noc)) +
  geom_bar(stat = "identity") +
  labs(title = "Top Sports with Highest Medal-Winning Countries", 
       x = "Sport", y = "Medal Count", fill = "Country") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


#Total Medals by Country and Sport in Selected Sports
```{r}
library(dplyr)
library(ggplot2)

selected_sports <- c("Swimming", "Gymnastics", "Wrestling", "Football", "Hockey", "Boxing", "Badminton", "Archery")
medals_by_country_sport <- olympics %>%
  filter(sport %in% selected_sports, !is.na(medal)) %>%
  group_by(noc, sport) %>%
  summarise(total_medals = n(), .groups = "drop") %>%
  arrange(desc(total_medals))
top_countries <- medals_by_country_sport %>%
  group_by(noc) %>%
  summarise(total_medals = sum(total_medals), .groups = "drop") %>%
  arrange(desc(total_medals)) %>%
  head(10) %>%
  pull(noc)
filtered_data <- medals_by_country_sport %>%
  filter(noc %in% top_countries)
ggplot(filtered_data, aes(x = noc, y = total_medals, fill = sport)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  labs(
    title = "Total Medals by Country and Sport in Selected Sports",
    x = "Country",
    y = "Total Medals",
    fill = "Sport"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_viridis_d()
```
